Embedded Scalars, Preferred Readings \& Intonation: {A}n Experimental Revisit

### 2014-09-01 text/paper/paper4.tex
K2 alternative explanation modern traditionalism

Whereas the existence of such implicatures is rather uncontroversial
for plain occurrences of \emph{some}, it is still an issue of debate
whether comparable enrichments are also found in so-called
``embedded'' cases, where the scalar item \emph{some} occurs in the
scope of other logical operators. This paper deals with to such cases.
In \as-sentences (mnemonic for \textit{all \dots some \dots}) as in
(\ref{bsp:AE}) the scalar item \emph{some} is embedded under universal
quantifier \emph{all}. In \es-sentences (mnemonic for \textit{exactly
  one \dots some \dots}) like (\ref{bsp:GE}) \emph{some} takes scope
under the non-monotonic quantifier \emph{exactly one}.  According to
current pragmatic theory, there are at least three relevant candidate
readings for \as- and \es-sentences: (i) a \emph{literal reading} like
in (\ref{bsp:AE-Literal}) and (\ref{bsp:GE-Literal}) where \emph{some}
has only its literal meaning; (ii) a \emph{global reading} like in
(\ref{bsp:AE-Global}) and (\ref{bsp:GE-Global}) where the meaning of
(\ref{bsp:AE}) is enriched and (\ref{bsp:GE}) with the negation of the
corresponding sentences (\ref{bsp:AE-Alternative}) and
(\ref{bsp:GE-Alternative}) respectively; and also (iii) a \emph{local
  reading} like in (\ref{bsp:AE-Local}) and (\ref{bsp:GE-Local}) where
\emph{some} is interpreted as \emph{some but not all} in the scope of
the embedding quantifier.

@@

% Section~\ref{sec:get-know-your} will discuss these readings
% in more detail, showing that all three readings of each sentence are
% distinct but logically dependent on each other in interesting ways.

\noindent Given this plurality of hypothesized readings, the question
arises:

\begin{exe}
\ex How salient or preferred are these theoretically conceivable
  readings compared to each other (if they are detectable by empirical
  means at all)?  
\end{exe}

\noindent Addressing this question empirically is relevant because it
lies at the heart of a current debate about the nature of quantity
inferences and the demarkation between semantics and pragmatics. The
controversy is usually seen as one between two main competing
theoretical positions. \emph{Traditionalism} holds in particular that
local readings should not occur frequently unless marked by special
intonation. \emph{Grammaticalism} holds that local readings are
unproblematically available and, at least under the most prominent of
selection principles for candidate readings, even at least as salient
as the literal reading (see Section~\ref{sec:grammaticalism}). To
distinguish between these opposing positions, it is therefore
necessary to assess the empirical observability and relative
preference of readings for \as- and \es-sentences under conditions
that control for prosodic markedness.

A number of empirical studies have already examined sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE})
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. Unfortunately,
results from these studies do not provide uncontroversial
evidence. This is due to several reasons. Firstly, previous studies
have only accumulated limited evidence pertaining to the preference
relations between candidate readings of both types of
sentences. Secondly, as \citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences
suggested by studies of \citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev}, might well be confounds of
what van Tiel and Geurts call \emph{typicality effects} (see
Section~\ref{sec:previous-studies} for discussion). Finally, and most
importantly, previous studies presented target sentences
visually. However, traditionalists admit to the availability of local
readings as long as the utterance is prosodically marked
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. We
will address this extant position as the \emph{prosodic markedness
  hypothesis}. If intonation plays a role, it is vital to present
target sentences auditorily, and to control for effects of ``silent
prosody'' that subjects may apply when reading a sentence
\citep[see][]{Bader98,Fodor98}.

This paper presents results form an experimental study designed to
deal with the above mentioned problems. We presented sentence
materials auditorily and explicitly manipulated intonational stress on
embedded scalar items. To cater for potential pictorial confounds,
such as from typicality, we employed an \emph{incremental verification
  task} in which parts of a to-be-evaluated picture were hidden and
only revealed piece-wise at the participants' request until they felt
able to give a binary truth-value judgement
\citep[see][]{Conroy2008}. This way we obtained categorical behavioral
data that is both more clearly indicative of the reading that subjects
assumed and mor independent of any potential pictorial confounds, such
as the mentioned typicality effects. Moreover, this method also gives
information about reading preferences. We validated this by including
control conditions with ambiguous sentences with a well-known
preference on available readings.

% The data obtained by this method is problematic for both
% traditionalism and grammaticalism. Contrary to the predictions of
% traditionalism, we observed many answers indicative of local readings
% of \es-sentences, irrespective of intonational markedness. But also
% grammaticalism is incongruent with the observed pattern in our
% data. Backed up by the strongest meaning hypothesis
% \citep{DalrympleKanazawa1998:Reciprocal-Expr}, grammaticalism predicts
% local readings to be the preferred readings, but we found no evidence
% for that. More strongly even, the preference order among putative
% readings suggested by our data cannot even in principle be explained
% by comparisons of logical strength among putative readings. In sum,
% our data sheds doubt on both traditionalism and grammaticalism in
% their currently most common forms. We conclude the paper by some
% speculation about how either position might be amended in the light of
% our data.

@@

Finally, we need to control for ``silent prosody.'' A growing number
of evidence suggests that numerous factors might influence accent
placement and prosodic phrasing even while reading, amongst them
default accentuation, constituent length, rhythmic phenomena, and
individual variation
\citep[e.g.][]{Augurzky08,Bader98,Fodor98,Fodor02,Kentner12,
  Steinhauer01}. If we adopt the prosodic markedness hypothesis
described in Section~\ref{sec:inton-mark-hypoth}, as many
traditionalists do, we predict that the availability of a local
reading hinges on the realization of contrastive stress on the scalar
item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}.
But if intonation can have this role, it is necessary to control for
silent accent placement. Ideally, therefore, we should present
sentences auditorily and systematically manipulate contrastive stress.

@@

The number of responses that fell out of our classification scheme
differs substantially between \as- and \es-conditions. There are
7.0\% / 4.4\% unclassifiable judgements in the neutral / accented
\as-conditions, but 19.3\% / 17.6\% in the neutral / accented
\es-conditions respectively. A look at the left-hand side of
Table~\ref{tab:count_data_CC} reveals that there is some systematicity
in 'error' answers. Most answers that we classified as 'errors' have a
plausible explanation: random errors, a handful of mistakes
resulting from confusing buttons for \emph{true} and \emph{false} and
a few 'spill-over errors' where participants gave responses one
position too late. There is one notable exception, though. A
substantial number of \emph{true} responses were given at the second
position of \es-sequences (9.7\% / 8.8\%) which is shown in
Figure~\ref{fig:exseqES2}. Although this position in the sequence is
the critical position to reveal a global reading, the latter would be
indicated by a \emph{false} response. It appears that subjects read
\es-sentences as if the non-monotonic quantifier \emph{exactly one}
was a monotonic \emph{at least one}, without pragmatic enrichment of
\emph{some}.\fn{If \emph{some} is pragmatically strengthened under the
  scope of \emph{at least one}, we would expect \emph{true} judgements
  at position 3. Although exactly 3 \emph{true} answers occurred there
  for each neutral and accented \es-conditions, the main point to
  notice is that this misreading of the non-monotonic quantifier does
  not explain \emph{true} judgements at position 5. These are
  therefore still indicative of local readings.} 16 out of the total
21 answers in this category come from 5 subjects, while the remaining
5 were single answers by 5 different subjects.
% subjects 1, 9, 12, 21, 84
This suggests that at least for some subjects non-monotonic
quantifiers were problematic. (We will come back to this point later.)

@@

High scores of success in control conditions show that participants
understood the incremental verification task and were able to give
judgements at the right point in a picture sequence. Results for the
preference-related controls show that our design is able to detect
multiple readings and preferences among these. Crucially, the amount
of response types for \lc- and \ec-readings reflected the known
preference for the former, with a mild bias for readings that can be
judged later in the sequence. Also, the incremental verification task
was generally sensitive to prosody. However, significant prosodic
effects were restricted to preference-related controls. Contrastively
accenting scalar \emph{some} generally led to more local readings, but
that effect was not significant. Local readings occurred in
\es-sentences, independently of prosody, but required prosodic marking
in \as-sentences. Our analyses support the impression that inspection
of the raw data gives, namely that the general qualitative pattern of
reading preferences is the same for \as- and \es-conditions: literal
readings are preferred over local readings which are preferred over
global readings. More concretely, our analyses suggest the following
preference/availability patterns (bracketed readings are unattested,
in the sense that they are either unavailable or heavily
dispreferred):

@@

Traditionalism easily explains the absence of global readings. Weak
traditionalism, introduced in Section~\ref{sec:traditionalism}, for
instance, would hold that the extra assumptions needed to derive
global readings (e.g., a speaker competence assumption) are not
available or only allow for epistemic weak implicatures that would
show as responses indicative of literal readings in our task. In
support of this view, traditionalists could argue that the task's
rather high processing demands lead to a suppression of global
readings, in line with the findings of
\citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. The small number of local readings in
\as-sentences with contrastive stress is easily explained by adopting
the prosodic markedness hypothesis. What traditionalism does not
predict is the high number of answers indicative of local readings in
the \es-conditions. Resorting to the prosodic markedness hypothesis
does not help, because local readings were observed for accented as
well as in unaccented \es-sentences. It would also remain unclear why
prosodic cues should have such differential effects on \as- and
\es-sentences.

Grammaticalism, on the other hand, fails to predict the pattern in
(\ref{bsp:observed-preferences}) entirely if it equates preference for
readings with logical strength. Grammaticalism predicts that literal
readings of \as-sentences are the least preferred, contrary to
(\ref{bsp:observed-preferences}). Again, we need to bear the caveat of
Section~\ref{sec:grammaticalism} in mind that we are only evaluating
grammaticalism with respect to a particular selection criterion
here. The verdict, however, seems to be clear. If
(\ref{bsp:observed-preferences}) is the actual preference order over
readings, then meaning selection in terms of logical strength is
generally on the wrong track, because no account that merely looks at
the logical strengths of readings can plausibly predict the preference
pattern in (\ref{bsp:observed-preferences}) for both \as- and
\es-sentences at the same time.

@@

Traditionalism, on the other hand, has to explain the possibility of
local readings of prosodically unmarked \es-sentences, and, perhaps,
also the large number of local readings in accented \es-sentences
compared to that in accented \as-sentences. To do so within a
traditionalist framework is difficult, if possible at all. A common
approach is to account for local readings as the result of some other
phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that the additional assumption of ``shallow
interpretation'' of the non-monotonic quantifiers might save
traditionalism.

As discussed in Section~\ref{sec:previous-studies},
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue that the
distribution of responses to \as-sentences found by
\citet{CliftonDube2010:Embedded-Implic}, as well as
\citet{ChemlaSpector2010:Experimental-Ev} could be explained by
differences in how \emph{typical} the presented pictures were for an
\as-sentence. Notice that this does not apply in the present situation
where the task is to explain away the local readings of \es-sentences,
for which no alternative explanation in terms of typicality has been
offered. Moreover, to apply typicality-explanations to the incremental
verification task, one would have to reason about the typicality of
parts of pictures or to figure in subjects' expectations of pictorial
typicality given their uncertainty about parts of the picture that
they have not yet seen. On intuitive grounds, we take either extension
to be implausible. But even if this line of explanation can be
plausibly extended to incrementally revealed picture verification, it
would still be unclear why typicality-effects should show so strongly
in categorical truth-value judgements, and why there is such a stark
contrast between \as- and \es-sentences.

@@

On the other hand, what might make this contrast-based explanation
more plausible is the observation that subjects had difficulties with
\es-sentences. As discussed in Section~\ref{sec:results}, there were a
non-negligible number of \emph{true} answer at step 2 in the
\es-sequence. This can be explained on the assumption that
participants read the non-monotonic quantifier \emph{exactly one} as a
monotonic quantifier \emph{at least one}. Similar ``shallow
interpretation'' of \emph{exactly one} would explain the high number
of local answers for \es-sentences (and for \es-sentence alone). If
participants processed the truth-value for an \es-sentence by looking
for a \emph{unique} element that could be singled out by a description
like ``\dots is connected to some of its \dots'', they would need to
unravel the \es-sequence up to the last picture where the relevant
visual contrast is stronger. A ``shallow interpretation''-theory also
sidesteps the puzzle why answers are given that are incompatible with
the proper semantic meaning of \es-sentences. Together, pictorial
contrast effects in symbiosis with a ``shallow interpretation'' of
non-monotonic quantifiers could be an alternative explanation of
(\ref{bsp:observed-preferences}) for traditionalism.

>>>>

Whereas the existence of such implicatures is rather uncontroversial
for plain occurrences of \emph{some}, it is still an issue of debate
whether comparable enrichments are also found in so-called
``embedded'' cases, where the scalar item \emph{some} occurs in the
scope of other logical operators. This paper deals with two such cases.
In \as-sentences (mnemonic for \textit{all \dots some \dots}) as in
(\ref{bsp:AE}) the scalar item \emph{some} is embedded under universal
quantifier \emph{all}. In \es-sentences (mnemonic for \textit{exactly
  one \dots some \dots}) like (\ref{bsp:GE}) \emph{some} takes scope
under the non-monotonic quantifier \emph{exactly one}.  According to
current pragmatic theory, there are at least three relevant candidate
readings for \as- and \es-sentences: (i) a \emph{literal reading} like
in (\ref{bsp:AE-Literal}) and (\ref{bsp:GE-Literal}) where \emph{some}
has only its literal meaning; (ii) a \emph{global reading} like in
(\ref{bsp:AE-Global}) and (\ref{bsp:GE-Global}) where the meaning of
(\ref{bsp:AE}) is enriched and (\ref{bsp:GE}) with the negation of the
corresponding sentences (\ref{bsp:AE-Alternative}) and
(\ref{bsp:GE-Alternative}) respectively; and also (iii) a \emph{local
  reading} like in (\ref{bsp:AE-Local}) and (\ref{bsp:GE-Local}) where
\emph{some} is interpreted as \emph{some but not all} in the scope of
the embedding quantifier.

@@

Given this plurality of hypothesized readings, the question arises:
how salient or preferred are these theoretically conceivable readings
compared to each other (if they are detectable by empirical means at
all)?  Addressing this question empirically is relevant because it
lies at the heart of a current debate about the nature of quantity
inferences and the demarkation between semantics and pragmatics. The
controversy is usually seen as one between two main competing
theoretical positions. \emph{Traditionalism} holds in particular that
local readings should not occur frequently unless marked by special
intonation. \emph{Grammaticalism} holds that local readings are
unproblematically available and, at least under the most prominent of
selection principles for candidate readings, even at least as salient
as the literal reading (see Section~\ref{sec:grammaticalism}). To
distinguish between these opposing positions, it is therefore
necessary to assess the relative salience of readings for \as- and
\es-sentences under conditions that control for prosodic markedness.

A number of empirical studies have already examined sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE})
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. Unfortunately,
results from these studies do not provide uncontroversial
evidence. This is due to several reasons. Firstly, previous studies
have only accumulated limited evidence pertaining to the relative
salience of candidate readings of both types of sentences. Secondly,
as \citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences
suggested by studies of \citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev}, might well be confounds of
what van Tiel and Geurts call \emph{typicality effects} (see
Section~\ref{sec:previous-studies} for discussion). Finally, and most
importantly, previous studies presented target sentences
visually. However, traditionalists acknowledge the availability of
local readings for prosodically marked utterances
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. We
will address this extant position as the \emph{prosodic markedness
  hypothesis} and argue that, if intonation is hypothesized to play a
central role, it is vital to present target sentences auditorily, and
to control for effects of ``silent prosody'' that subjects may apply
when reading a sentence \citep[see][]{Bader98,Fodor98}.

This paper presents results form an experimental study designed to
deal with the above mentioned problems. We presented sentence
materials auditorily and explicitly manipulated intonational stress on
embedded scalar items. To cater for potential pictorial confounds,
such as from typicality, we employed an \emph{incremental verification
  task} in which parts of a to-be-evaluated picture were hidden and
only revealed piece-wise at the participants' request until they felt
able to give a binary truth-value judgement
\citep[see][]{Conroy2008}. This way we obtained categorical behavioral
data that is both more clearly indicative of the reading that subjects
assumed and more independent of any potential pictorial confounds, such
as the mentioned typicality effects. Moreover, this method also gives
information about reading preferences. We validated this by including
control conditions with ambiguous sentences with a well-known
preference on available readings.

The data obtained by this method is \emph{prima facie} problematic for
both traditionalism and grammaticalism. Contrary to the predictions of
traditionalism, we observed many answers indicative of local readings
of \es-sentences, irrespective of intonational markedness. But also
grammaticalism is incongruent with the observed pattern in our
data. Backed up by the strongest meaning hypothesis
\citep{DalrympleKanazawa1998:Reciprocal-Expr} as a disambiguation
criterion \citep[as suggested by][]{ChierchiaFox2008:The-Grammatical},
grammaticalism predicts local readings to be the preferred readings,
but we found no evidence for that. More strongly even, the preference
order among putative readings suggested by our data cannot even in
principle be explained by comparisons of logical strength among
putative readings. In sum, our data sheds doubt on both traditionalism
and grammaticalism in their currently most common forms. We conclude
the paper by some speculation about how either position might be
amended in the light of our data.

@@

Finally, we need to control for ``silent prosody.'' Several studies
suggest that numerous factors might influence accent placement and
prosodic phrasing even while reading, amongst them default
accentuation, constituent length, rhythmic phenomena, and individual
variation
\citep[e.g.][]{Bader98,Fodor98,Steinhauer01,Fodor02,Augurzky08,BreenWatson2011:Intonational-ph,Kentner12}. If
we adopt the prosodic markedness hypothesis described in
Section~\ref{sec:inton-mark-hypoth}, as many traditionalists do, we
predict that the availability of a local reading hinges on the
realization of contrastive stress on the scalar item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}.
But if intonation can have this role, it is necessary to control for
silent accent placement. Ideally, therefore, we should present
sentences auditorily and systematically manipulate contrastive stress.

@@

The number of responses that fell out of our classification scheme
differs substantially between \as- and \es-conditions. There are 7.0\%
/ 4.4\% unclassifiable judgements in the neutral / accented
\as-conditions, but 19.3\% / 17.6\% in the neutral / accented
\es-conditions respectively. A look at the left-hand side of
Table~\ref{tab:count_data_CC} reveals that there is some systematicity
in 'error' answers. Most answers that we classified as 'errors' have a
plausible explanation: random errors, a handful of mistakes resulting
from confusing buttons for \emph{true} and \emph{false} and a few
'spill-over errors' where participants gave responses one position too
late. There is one notable exception, though. A substantial number of
\emph{true} responses were given at the second position of
\es-sequences (9.7\% / 8.8\%) which is shown in
Figure~\ref{fig:exseqES2}. Although this position in the sequence is
the critical position to reveal a global reading, the latter would be
indicated by a \emph{false} response. It appears that subjects read
\es-sentences as if the non-monotonic quantifier \emph{exactly one}
was a monotonic \emph{at least one}, without pragmatic enrichment of
\emph{some}. 16 out of the total 21 answers in this category come from
5 subjects, while the remaining 5 were single answers by 5 different
subjects.
% subjects 1, 9, 12, 21, 84
This suggests that at least for some subjects non-monotonic
quantifiers were problematic. (We will come back to this intriguing
observation later in Section~\ref{sec:discussion-1}, where we suggest
that a non-monotonic reading of \emph{exactly one} might also explain
the high number of \emph{true} answers in step 5 of the
\es-conditions.)

@@

High scores of success in control conditions show that participants
understood the incremental verification task and were able to give
judgements at the right point in a picture sequence. Results for the
preference-related controls show that our design is able to detect
multiple readings and preferences among these. Crucially, the amount
of response types for \lc- and \ec-readings reflected the known
preference for the former, with a mild bias for readings that can be
judged later in the sequence. Also, the incremental verification task
was generally sensitive to prosody. However, significant prosodic
effects were restricted to preference-related controls. Contrastively
accenting scalar \emph{some} generally led to more local readings, but
that effect was not significant. Local readings occurred in
\es-sentences, independently of prosody, but required prosodic marking
in \as-sentences. Our analyses support the impression that inspection
of the raw data gives, namely that the general qualitative pattern of
reading preferences is the same for \as- and \es-conditions: literal
readings are preferred over local readings which are preferred over
global readings. More concretely, our analyses suggest the following
preference patterns (bracketed readings are not supported by our data
and the previous analyses):

@@

Traditionalism easily explains the absence of global readings. Weak
traditionalism, introduced in Section~\ref{sec:traditionalism}, for
instance, would hold that the extra assumptions needed to derive
global readings (e.g., a speaker competence assumption) are not
available or only allow for epistemic weak implicatures that would
show as responses indicative of literal readings in our task. In
support of this view, traditionalists could argue that the task's
rather high processing demands lead to a suppression of global
readings, in line with the findings of
\citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. The small number of local readings in
\as-sentences with contrastive stress is easily explained by adopting
the prosodic markedness hypothesis. What traditionalism does not
predict is the high number of answers indicative of local readings in
the \es-conditions. Resorting to the prosodic markedness hypothesis
does not help, because local readings were observed for accented as
well as for unaccented \es-sentences. It would also remain unclear why
prosodic cues should have such differential effects on \as- and
\es-sentences.

Grammaticalism, on the other hand, fails to predict the pattern in
(\ref{bsp:observed-preferences}) entirely if it equates preference for
readings with logical strength. Grammaticalism predicts that literal
readings of \as-sentences are the least preferred, contrary to
(\ref{bsp:observed-preferences}). Again, we need to bear the caveat of
Section~\ref{sec:grammaticalism} in mind that we are only evaluating
grammaticalism with respect to a particular disambiguation criterion
here. The verdict, however, seems to be clear. If
(\ref{bsp:observed-preferences}) is the actual preference order over
readings, then meaning disambiguation in terms of logical strength is
generally on the wrong track, because no account that merely looks at
the logical strengths of readings can plausibly predict the preference
pattern in (\ref{bsp:observed-preferences}) for both \as- and
\es-sentences at the same time.

@@

Traditionalism, on the other hand, has to explain the possibility of
local readings of prosodically unmarked \es-sentences, and, perhaps,
also the large number of local readings in accented \es-sentences
compared to the much lower number in accented \as-sentences. To do so
within a strict traditionalist framework seems difficult. A first
options might be to account for local readings as the result of some
other phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that a ``modern traditionalist'' explanation of the
answer pattern might be available if we allow for the possibility that
\emph{exactly one} gets an unexpected ``referential interpretation.''

As discussed in Section~\ref{sec:previous-studies},
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue that the
distribution of responses to \as-sentences found by
\citet{CliftonDube2010:Embedded-Implic}, as well as
\citet{ChemlaSpector2010:Experimental-Ev} could be explained by
differences in how \emph{typical} the presented pictures were for an
\as-sentence. Notice that this does not apply in the present situation
where the task is to explain away the local readings of \es-sentences,
for which no alternative explanation in terms of typicality has been
offered. Moreover, to apply typicality-based explanations to the
incremental verification task, one would have to reason about the
typicality of parts of pictures or to figure in subjects' expectations
of pictorial typicality given their uncertainty about parts of the
picture that they have not yet seen. On intuitive grounds, we take
either extension to be implausible. But even if this line of
explanation can be plausibly extended to incrementally revealed
picture verification, it would still be unclear why typicality-effects
should show so strongly in categorical truth-value judgements, and why
there is such a stark contrast between \as- and \es-sentences.

@@

On the other hand, there is a conceivable ``modern traditionalist''
alternative explanation of the puzzling data for \es-sentences that
would also account for the large and systematic error
responses. Recall from Section~\ref{sec:results} and
Table~\ref{tab:count_data_CC} on \pageref{tab:count_data_CC} that a
surprisingly high number of \emph{true} answers was given at the
second step of the \es-sequence. This answer does not correspond to
any of the three candidate readings that the previous literature has
focused on. We hypothesized in Section~\ref{sec:results} that these
readings can be explained under the assumption that \emph{exactly one}
gets a reading similar to \emph{there is (at least) one}, while the
scalar item \emph{some} would just receive its semantic
interpretation. Such a reading of \emph{exactly one}, although
surprising, could actually be supported by some theories of numerals
and modifiers
\citep[e.g.][]{Geurts2006:Take-five,MartyChemla2014:Phantom-reading},
as the outcome of an ``existential closure'' type-shifting rule in the
sense of \citet{Partee1986:Noun-phrase-int} to the effect that
\emph{exactly one} is mapped onto an existential reading of the form
``there is a group with cardinality exactly one.''

What happens if \emph{some} is additionally strengthened in some way
or other under such a reading of \emph{exactly one}? If \emph{some} is
pragmatically strengthened under the scope of \emph{at least one} in a
pure traditionalist manner by adding the negation of a sentence type
\emph{at least one \dots all \dots}, we would expect \emph{true}
judgements at position 3 of the \es-sequence. Although exactly 3
\emph{true} answers occurred there for each neutral and accented
\es-conditions respectively, the main point to notice is that this
construal does not explain \emph{true} judgements at position 5. These
would therefore still seem indicative of local readings. However,
under the hypothesized existentially closed reading of \emph{exactly
  one}, a ``modern traditionalist'' approach that takes discourse
effects into account \citep[][Chapter~8.4]{Geurts2010:Quantity-Implic}
would predict the availability of a pragmatically strengthened reading
of \es-sentences of the form: ``there is a group of cardinality exactly
one and that group has the property that its members are connected to
some but not all of their surrounding elements.'' This would explain
the observed patterns in \es-conditions. Participants who read
\emph{exactly one} as a non-monotonic quantifier would, pace
traditionalist logic, prefer a literal interpretation and answer
\emph{false} at step 3 unless they made a mistake. Participants who
assumed an existentially-closed quantifier reading could, according to
a discourse-sensitive ``modern traditionalism'', apply a literal or a
pragmatically strengthened reading. In the former case, they would
answer \emph{true} at step 2, in the latter case they would answer
\emph{true} at step 5, without this being necessarily a ``local
reading'' in the sense put forward by proponents of grammaticalism.

This alternative explanation is not without problems and is not backed
up by a full account of discourse-based pragmatic reasoning. In
conclusion, then, it appears that our data provides interesting
challenges to both traditionalism and grammaticalism.
